### 2023-08-03 analysis/analysis.Rmd
Switch to stars from raster/terra

---
title: "Palaeodistribution of crop progenitors and wild foods in the Southern Levant"
author: "Joe Roe"
format: html
---

```{r dependencies}
library("dplyr")
library("fs")
library("ggplot2")
library("ggspatial")
library("here")
library("parsnip")
library("purrr")
library("ranger")
library("rpaleoclim")
library("sf")
library("stars")
library("stringr")
library("tibble")
library("tidyr")

library("BadiaPaleoFloraENM") # this package, install with devtools::install()
```

```{r definitions}
latlong <- 4326
utm37n <- 32637

w_asia <- st_bbox(c(xmin = 30, xmax = 50, ymin = 25, ymax = 40), crs = 4326)
s_levant <- st_bbox(c(xmin = 34, xmax = 39, ymin = 29, ymax = 34), crs = 4326)
```

## Introduction

## Background

## Methods and materials

```{r data-basemap}
# Turn off s2 while we deal with Natural Earth's dodgy geometries
use_s2 <- sf_use_s2(FALSE)

ne_countries <- read_sf(raw_data("ne"), "ne_10m_admin_0_countries")

ne_lakes <- read_sf(raw_data("ne"), "ne_10m_lakes") |> 
  st_geometry() |>
  st_crop(buffer_bbox(w_asia, 10)) |>
  st_union()

ne_ocean <- read_sf(raw_data("ne"), "ne_10m_ocean") |>
  st_geometry() |>
  st_crop(buffer_bbox(w_asia, 10)) |>
  st_union()

ne_land <- read_sf(raw_data("ne"), "ne_10m_land") |> 
  st_geometry() |>
  st_crop(buffer_bbox(w_asia, 10)) |>
  st_difference(ne_ocean) |>
  st_difference(ne_lakes) |>
  st_union()

s_levant_land <- st_intersection(st_as_sfc(s_levant), ne_land)

# Restore previous setting of s2
sf_use_s2(use_s2)
```

### Occurrence data

Occurrence data was obtained from GBIF.

```{r data-occ}
gbif_data <- function(scientificName, ...) {
  gbif_data <- rgbif::occ_data(scientificName = scientificName, ...)
  gbif_data$data
}

gbif_path <- here::here("analysis", "data", "derived_data", "gbif")

taxa <- c("Hordeum spontaneum" = "Hordeum spontaneum")
          # "Bolboshoenus spp." = "Bolboschoenus")

# Get occurrence data from GBIF
# Including only records with coordinates and no recorded issues with geospatial
#   data.
taxa |>
  map(gbif_data, hasCoordinate = TRUE, hasGeospatialIssue = FALSE, 
      geometry = w_asia, limit = 100000) |>
  list_rbind(names_to = "taxon") ->
  occ

# Exclude:
# * Coordinate uncertainty of 5 km or more (≈2.5 mins)
# * Fossil occurrences
occ |>
  filter(coordinateUncertaintyInMeters < 5000 | is.na(coordinateUncertaintyInMeters)) |>
  filter(basisOfRecord != c("FOSSIL_SPECIMEN")) ->
  occ

# Remove duplicated coordinates (as recommended by CITE)
occ |>
  group_by(taxon) |>
  distinct(decimalLongitude, decimalLatitude, .keep_all = TRUE) |>
  ungroup() ->
  occ

# Convert to sf
occ |>
  st_as_sf(coords = c("decimalLongitude", "decimalLatitude"), 
           crs = 4326, remove = FALSE) ->
  occ
```

```{r plot-occ-distribution}
ggplot(occ) + 
  facet_wrap(vars(taxon)) +
  layer_spatial(ne_land, fill = "#ffffff") +
  annotation_spatial(s_levant, fill = alpha("red", 0.15), colour = "red") +
  layer_spatial(occ, alpha = 0.1, symbol = 16) +
  layer_spatial(ne_countries, fill = NA) +
  coord_sf(xlim = w_asia[c("xmin", "xmax")], 
           ylim = w_asia[c("ymin", "ymax")],
           crs = utm37n,
           default_crs = latlong)
```

<!-- TODO: following @ValaviEtAl2022, more accurate to call pseudo-absences
  "background samples" -->
Occurrence data only tells us where a species is present;
there is no definitive information on where the species is *not* found.
We therefore need to generate random background points or "pseudo-absences" to feed to the model.
There are several ways to do this. 
We follow the advice of Barbet-Massin et al. (2012) for regression-based species distribution models and use a large (~10000) random sample of points, weighted equally against the presences in the regression.
@ValaviEtAl2022 recommends using a very large background sample for random forest models.

```{r psabsences, include=FALSE}
n_background <- 1000

occ |>
  mutate(present = TRUE) |>
  bind_rows(background_sample(s_levant_land, 
                              n_background,
                              coord_x = "decimalLongitude",
                              coord_y = "decimalLatitude", 
                              taxon = "Hordeum spontaneum", 
                              present = FALSE)) ->
  occ

# Weights
#occ %>% 
#  group_by(taxon, occurrence) %>% 
#  mutate(weight = 1 + (n_psab / n() - 1)) ->
#  occ
```

### Climate data

```{r data-climate}
raw_data("paleoclim") |>
  dir_ls() |>
  map(dir_ls, glob = "*.tif") |>
  map(read_stars) |>
  map(\(x) set_names(x, path_ext_remove(path_file(names(x))))) ->
  climate

# Tidy up names
names(climate) |>
  basename() |>
  str_remove(regex("chelsa_", ignore_case = TRUE)) |>
  str_split_i("_", 1) ->
  names(climate)
```

```{r training-climate}
occ |>
  mutate(bio = extract_points(climate$cur, occ$geometry)) |>
  unpack(bio) ->
  occ
```

### Other predictor data

Soil, terrain (incl. TWI, slope?)

### Random Forest

```{r model}
# TODO: Investigate where these NAs are coming from
rand_forest(mode = "regression") |>
  fit(as.numeric(present) ~ ., data = select(occ, present, starts_with("bio_"))) ->
  rf_fit
```

## Results

### Modelled ecological niches

* On current data
* Sensitivity to climate fluctuations

### Hindcasting

```{r predict}
# TODO: deal with missing data in pred vars
# TODO: turn back to something spatial? stars?
predict(rf_fit, climate$cur)
```

### Verification

## Discussion