### 2025-01-08 analysis/appendix.qmd
Revise species list

---
title: "Biogeography of crop progenitors and wild plant resources in the terminal Pleistocene and Early Holocene of West Asia, 14.7–8.3 ka"
subtitle: "Fitted model summaries and predicted palaeodistributions"
author: 
- name: Joe Roe
  id: jr
  orcid: 0000-0002-1011-1244
  email: joeroe@hey.com
  affiliations:
    - name: University of Bern
      country: Switzerland
    - name: University of Copenhagen
      country: Denmark
  corresponding: true
- name: Amaia Arranz-Otaegui
  id: aao
  orcid: 0000-0002-5091-6426
  affiliation: 
    - name: University of the Basque Country
      country: Spain
abstract: |
  This appendix describes ecological niche models of 81<!-- TODO keep updated -->
  plant species relevant to the subsistence of Late Epipalaeolithic (15–11.7 ka) 
  and Neolithic (11.7–8.2 ka) societies in West Asia. It includes summaries of 
  the training data and performance metrics of the fitted models together with 
  predicted palaeodistributions of each species for three simulated 
  climatologies over the Pleistocene/Holocene boundary (the Bølling–Allerød, 
  14.7–12.9 ka; Younger Dryas, 12.9–11.7 ka; and Early Holocene, 11.7–8.3 ka) as
  well as current conditions.
  For a full description of the methodology, references, and discussion of the 
  results, please see the main text. The data and R code used to produce the
  models, as well as the full predicted palaeodistributions in raster format,
  are achived with Zenodo at <https://doi.org/> <!-- TODO insert Zenodo DOI -->
format:
  html:
    toc: true
execute:
  echo: false
---

```{r dependencies, message=FALSE}
library("cowplot")
library("dplyr")
library("here")
library("forcats")
library("ggplot2")
library("ggspatial")
library("gt")
library("khroma")
library("patchwork")
library("purrr")
library("readr")
library("sf")
library("stars")
library("tidyr")
library("tune")
library("yardstick")
library("vip")

library("BadiaPaleoFloraENM") # this package, install with devtools::install()

options(knitr.duplicate.label = "allow")
```

```{r data, message=FALSE}
# Model
flora <- readRDS(derived_data("flora_niche_predictions.Rdata"))
archaeo_flora <- read_tsv(
  derived_data("archaeo_flora_predictions.tsv"),
  col_types = cols(
    source = col_character(),
    source_id = col_character(),
    source_site_name = col_character(),
    site_name = col_character(),
    latitude = col_double(),
    longitude = col_double(),
    phase = col_character(),
    phase_description = col_character(),
    phase_code = col_character(),
    age_start = col_double(),
    age_end = col_double(),
    taxon_source = col_character(),
    n = col_double(),
    prop = col_double(),
    reference = col_character(),
    taxon = col_character(),
    taxon_group = col_character(),
    genus = col_character(),
    family = col_character(),
    category = col_character(),
    founder_crop = col_character(),
    edibility = col_character(),
    grass_type = col_character(),
    legume_type = col_character(),
    age_mid = col_double(),
    period = col_character(),
    climate_period = col_character(),
    pred_present = col_logical(),
    pred_absent = col_logical()
  )
)

# Chronology
climate_periods <- tibble(
  period = c("Bølling-Allerød", "Younger Dryas", "Early Holocene", "Current"),
  code = c("ba", "yds", "eh", "cur"),
  date_range = c("(14.7-12.9 ka)", "(12.9-11.7 ka)", "(11.7-8.3 ka)", ""),
  label = fct_inorder(paste(toupper(code), date_range)),
  long_label = fct_inorder(paste(period, date_range)),
  start_bp = c(14700, 12900, 11700, NA),
  end_bp = c(12900, 11700, 8300, NA)
)

# Map
latlong <- 4326
w_asia_albers <- "+proj=aea +lat_1=22.5 +lat_2=42.5 +lon_0=40"

w_asia <- st_bbox(c(xmin = 25, xmax = 55, ymin = 22.5, ymax = 42.5), crs = 4326)

# Turn off s2 while we deal with Natural Earth's funky geometries
use_s2 <- sf_use_s2(FALSE)

ne_countries <- read_sf(raw_data("ne"), "ne_10m_admin_0_countries")

ne_rivers <- read_sf(raw_data("ne"), "ne_10m_rivers_lake_centerlines_scale_rank")

ne_ocean <- read_sf(raw_data("ne"), "ne_10m_ocean") |>
  st_geometry() |>
  st_crop(buffer_bbox(w_asia, 10)) |>
  st_union()

ne_land <- read_sf(raw_data("ne"), "ne_10m_land") |> 
  st_geometry() |>
  st_crop(buffer_bbox(w_asia, 10)) |>
  st_difference(ne_ocean) |>
  st_union()

# Restore previous setting of s2
sf_use_s2(use_s2)
```

```{r content, message=FALSE, results="asis"}
flora <- arrange(flora, taxon)
content <- pmap(
  flora, 
  function(taxon, taxon_detail, occ, niche, palaeodist, ...) {
    archaeo <- archaeo_flora[archaeo_flora$taxon == taxon,]
    knitr::knit_child(
      here("analysis", "appendix_section.qmd"),
      envir = environment(),
      quiet = TRUE
    )
  },
  archaeo_flora = archaeo_flora, 
)

cat(unlist(content), sep = "\n")
```