### 2025-08-14 analysis/appendix_section.qmd

# *`r taxon`*

```{r taxon-detail, eval=(length(taxon_aliases) > 0), results = "asis"}
cat("Including ", knitr::combine_words(taxon_aliases), ".", sep = "")
```

```{r fig-model, fig.cap=paste0("Fitted model summary for *", taxon, "*")}
fig_occ <- occ |>
  filter(as.logical(present)) |>
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326) |>
  ggplot() +
  annotation_spatial(ne_land, fill = "#ffffff") +
  annotation_spatial(ne_rivers, aes(linewidth = strokeweig),
                     colour = "lightblue") +
  geom_sf(size = 0.5) +
  layer_spatial(w_asia, fill = NA) +
  scale_linewidth_identity() +
  coord_sf(xlim = w_asia[c(1,3)], ylim = w_asia[c(2,4)],
           crs = w_asia_albers, default_crs = latlong) +
  labs(
    title = "GBIF occurrences",
    subtitle = paste0("N=", nrow(occ)),
    caption = "NB. occurrences outside of prediction region not shown"
  )

fig_roc_curve <- niche |>
  collect_predictions() |>
  roc_curve(.pred_TRUE, truth = present, event_level = "second") |>
  ggplot() +
  geom_abline(intercept = c(0, 0), slope = 1, linetype = "dashed") +
  geom_line(aes(1 - specificity, sensitivity)) +
  labs(title = "ROC curve") +
  theme_minimal_grid() +
  theme(
    plot.title = element_text(size = 12, face = "plain"),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 10)
  )

fig_vip <- niche |>
  extract_fit_engine() |>
  vip() +
  labs(
    title = "Variable importance",
    subtitle = "Top 10 (impurity)"
  ) +
  theme_minimal_grid() +
  theme(
    plot.title = element_text(size = 12, face = "plain"),
    plot.subtitle = element_text(size = 10),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 10)
  )

fig_occ + (fig_roc_curve / fig_vip) +
  plot_layout(widths = c(3, 1))
```

```{r fig-palaeodist, warning=FALSE, fig.height=16, fig.cap=paste0("Predicted palaeodistributions of *", taxon, "*")}
if (length(archaeo_pred) > 0) {
  archaeo_geom <- geom_sf(
    aes(shape = .pred_class),
    data = archaeo_pred |>
      mutate(period = climate_period) |>
      distinct(longitude, latitude, period, .keep_all = TRUE) |>
      st_as_sf(coords = c("longitude", "latitude"), crs = latlong)
  )
} else {
  archaeo_geom <- NULL
}

ggplot() +
  facet_wrap(
    vars(
      factor(period, levels = climate_periods$code, 
             labels = climate_periods$long_label)
    ),
    ncol = 1
  ) +
  annotation_spatial(ne_land, fill = "white") +
  geom_stars(
    data = select(palaeodist, prediction.TRUE.), 
    na.action = na.omit
  ) +
  archaeo_geom +
  annotation_spatial(ne_land, fill = NA) +
  annotation_spatial(w_asia, fill = NA) +
  scale_fill_batlow(reverse = TRUE) +
  scale_shape_manual(
    name = "Archaeological\noccurrences", 
    values = c(1, 16),
    labels = c("Not predicted at\nP(present)>0.5", "Predicted at\nP(present)>0.5")
  ) +
  coord_sf(crs = w_asia_albers) +
  labs(fill = "P(present)", x = NULL, y = NULL) +
  theme(
    strip.background = element_blank(),
    strip.text.y = element_text(face = "italic")
  )
```

```{r tbl-archaeo, eval=(length(archaeo_pred) > 0)}
archaeo_pred |>
  mutate(taxon = taxon) |>
  summarise(
    n_assemb = n(),
    mean_p = mean(prop),
    age_start = max(age_start),
    age_end = min(age_end),
    pred_present = all(as.logical(.pred_class)),
    .by = c(source, climate_period, site_name)
  ) |>
  mutate(
    climate_period = map(climate_period, function(x) {
      climate_periods[climate_periods$code == x,]$long_label[[1]]
    }),
    climate_period = unlist(climate_period),
    age_start = age_start / 1000,
    age_end = age_end / 1000
  )|>
  group_by(climate_period) |>
  arrange(desc(age_start)) |>
  select(climate_period, site_name, age_start, age_end, n_assemb, mean_p, source,
         pred_present) |>
  gt(caption = md(paste0("Archaeological occurrences of *", taxon, "*"))) |>
  cols_label(site_name = "Site", n_assemb = "N assemblages",
             mean_p = "Average prop.", source = "Source",
             age_start = "Age range", pred_present = "Predicted?") |>
  cols_merge(c(age_start, age_end), pattern = "{1}–{2} ka") |>
  fmt_number(c(age_start, age_end), decimals = 1) |>
  fmt_percent(c(mean_p)) |>
  fmt_tf(c(pred_present), tf_style = "yes-no") |>
  sub_missing()
```