### 2025-08-15 analysis/paper.tex

% Options for packages loaded elsewhere
% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  authoryear,
  preprint]{elsarticle}
\usepackage{xcolor}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{5}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother

\setlength{\emergencystretch}{3em} % prevent overfull lines

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage[]{natbib}
\bibliographystyle{elsarticle-harv}

\usepackage{booktabs}
\usepackage{caption}
\usepackage{longtable}
\usepackage{colortbl}
\usepackage{array}
\usepackage{anyfontsize}
\usepackage{multirow}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\journal{Open Quaternary}
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Biogeography of crop progenitors and wild plant resources in the terminal Pleistocene and Early Holocene of West Asia, 14.7--8.3 ka},
  pdfauthor={Joe Roe; Amaia Arranz-Otaegui},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\setlength{\parindent}{6pt}
\begin{document}

\begin{frontmatter}
\title{Biogeography of crop progenitors and wild plant resources in the
terminal Pleistocene and Early Holocene of West Asia, 14.7--8.3 ka}
\author[1,2]{Joe Roe%
\corref{cor1}%
}
 \ead{jar@hum.ku.dk} 
\author[3]{Amaia Arranz-Otaegui%
%
}

\affiliation[1]{organization={University of
Bern},country={Switzerland},countrysep={,},postcodesep={}}
\affiliation[2]{organization={University of
Copenhagen},country={Denmark},countrysep={,},postcodesep={}}
\affiliation[3]{organization={University of the Basque
Country},country={Spain},countrysep={,},postcodesep={}}

\cortext[cor1]{Corresponding author}

\begin{abstract}
This paper presents the first continuous, spatially-explicit
reconstructions of the palaeodistributions of 65 plant species found
regularly in association with early agricultural archaeological sites in
West Asia, including the progenitors of the first crops. We used machine
learning to train an ecological niche model of each species based on its
present-day distribution in relation to climate and environmental
variables. Predictions of the potential niches of these species at key
stages of the Pleistocene--Holocene transition could then be derived
from these models using downsampled data from palaeoclimate simulations.
Our models performed well against independent contemporary test data,
but their ability to predict the occurrence of specific species at
archaeological sites was much more variable, probably reflecting a
tendency of the method to underestimate the species' fundamental niche.
Nevertheless, the majority of species are predicted to have had more
restricted geographic distributions under past climate conditions
compared to today. Crop progenitors and several wild food species
modelled are modelled to have been concentrated in the Levant and, to a
lesser extent, Cyprus and Western Anatolia, suggesting these regions may
have served as glacial \emph{refugia}. The average size of species'
niche shrunk by an average of c.~25\% from the terminal Pleistocene to
the Early Holocene, indicating that economically significant plants were
adapted to cryo-arid conditions and did not, as often assumed, initially
respond positively to the `ameliorated' climate of the Holocene.
\end{abstract}

\end{frontmatter}

\section{Introduction}\label{introduction}

The Pleistocene--Holocene transition in West Asia marked a turning point
in global environmental history, as humans brought the first plants
under cultivation and began modifying surrounding ecosystems to support
their own subsistence. West Asia is part of the native range of a
remarkable number of domesticable plant species, including wild
relatives of wheat, barley, peas, lentils, and other crops of global
importance \citep{HarlanZohary1966, Diamond2002, ZoharyEtAl2012}. These
species supported uniquely dense and complex Late Epipalaeolithic
(15--11.7 ka) societies \citep{BarYosef1998, MaherEtAl2012} based on
foraging \citep{HarrisHillman1989, Colledge2001, WeissEtAl2004} and
eventually plant management and pre-domestication cultivation
\citep{Colledge2001, WeissEtAl2006, Harris2007, WillcoxEtAl2008}. The
first agro-ecosystems emerged as these plants were domesticated in the
Pre-Pottery Neolithic period (11.7--8.5 ka) and were shaped by the
broader ecosystem in which it was embedded.

Decades of research in archaeobotany and zooarchaeology have
reconstructed the subsistence economies of Late Epipalaeolithic and
Neolithic sites in great detail. Together with studies of other
environmental archaeological records and a variety of palaeoclimate
archives \citep{JonesEtAl2019}, they also tell us much about the
environments surrounding these settlements. However, each of these
sources of evidence is subject to the wide variety of taphonomic and
recovery biases that are inherent in any direct record of the past. They
are also, by definition, records of the (human) environment at
particular times and places. Interpolating these snapshots to give a
holistic picture of the regional ecologies is not straightforward -- to
date, it has tended to rely on non-explicit, inductive modelling. The
majority are also filtered through human action, producing a mixed
single that makes it difficult to disentangle anthropic effects from the
background of environmental change in this period of rapid climatic
alteration.

In this paper we present a complementary, deductive approach based on
ecological niche modelling. Rather than inferring environmental
conditions from preserved physical evidence, we predict the potential
niche of individual species relevant to human subsistence based on a
model of their current environmental niche and simulations of past
palaeoclimate. Though hypothetical, this gives us an independent line of
evidence on past ecologies that is independent of the environmental
archaeological and palaeoclimatic records. This means that the ancient
data can be reserved for assessing the model's ability to `hindcast'
past conditions. In this sense, discrepencies between the two records
are perhaps the most interesting result, as they indicate processes
affecting one or both records that are not fully accounted for and
therefore generate new questions. Our computational approach is also
readily scaled up, allowing us to model spatially-explicit
palaeodistributions for a large number of species, for the whole region,
under multiple reconstructed climate scenarios.

\section{Background}\label{sec-bg}

The transition to agriculture represents one of the most fundamental
changes in human history. West Asia is one of the regions where this
process has been studied in the most detail: decades of research have
traced the gradual development of a Neolithic way of life and the
changes that occurred in the plant species and their geographical
distribution as a result. Although archaeobotanical assemblages can be
biased due to issues of preservation, sampling, recovery techniques, and
lab procedures \citep{Dennel1976, HastorfPopper1988}---and although they
include not just food remains but plant resources that were used for
other purposes or arrived at the site accidentally
\citep{HastorfPopper1988}---large-scale studies have still revealed
coherent patterns in the exploitation of plants over time
\citep{ColledgeEtAl2004, ArranzOtaeguiEtAl2016}.

The possibility of an abrupt, geographically-constrained process of
plant domestication was proposed in the 1990s
\citep{HillmanDavies1990, HillmanDavies1992, HeunEtAl1997, OzkanEtAl2002}
and developed as an explanatory model in the 2000s
\citep{LevYadunEtAl2000, GopherEtAl2001, AbboEtAl2005}. As part of this
model, some authors
\citep{LevYadunEtAl2000, GopherEtAl2001, AbboEtAl2010, AbboEtAl2012}
argued that eight plant species, collectively referred to as `founder
crops' or the `Neolithic crop package' \citep{ZoharyHopf1988} were
selected and domesticated once, without any phase of pre-domestication
cultivation \citep[p.~177]{AbboEtAl2011}. This process could have been
rapid under strong artificial selection
\citep{HillmanDavies1990, HillmanDavies1992} and may have occurred in a
single region or `core area' -- generally located in southeast Turkey
\citep{LadizinskyAdler1976, HeunEtAl1997, OzkanEtAl2002, OzkanEtAl2005, MoriEtAl2003, LuoEtAl2007}.
From this single point of origin, it was supposed that domesticated or
semi-domesticated plants radiated outwards to other regions
\citep{AbboEtAl2006, KilianEtAl2007, OzkanEtAl2011}.

The `short gestation' paradigm was challenged by others
\citep{Helbaek1969, Harris1989, Kislev1989, Colledge2001, WeissEtAl2004, WillcoxEtAl2008, FullerEtAl2018}.
Helbæk \citep[in][]{Kirkbride1966} argued that before the appearance of
domesticated plants, a phase of cultivation of wild seeds must have
taken place. The existence of a phase of cultivation of morphologically
wild cereals or `pre-domestication cultivation' was identified in the
archaeological record through the study of plant domestication traits
such as grain size, shattering v. non-shattering rachises. The
archaeobotanical evidence shows that during the Pre-Pottery Neolithic A
(PPNA) cereals exhibited sizes similar to those recorded in domestic
species, but their dispersal mechanism was still the same as the one
present in morphologically-wild species \citep[i.e.~shattering,
see][]{Kirkbride1966, Kislev1989, HillmanEtAl2001, Colledge2001, WillcoxEtAl2008}.
This evidence suggested that wild cereal stands could have been
cultivated for as much as a thousand years before non-shattering
domestic forms became prevalent in the archaeological record
\citep{TannoWillcox2006, TannoWillcox2012, ArranzOtaeguiEtAl2016}.
Additional archaeobotanical
\citep{Colledge2001, WillcoxEtAl2008, WillcoxEtAl2009, RiehlEtAl2013, ArranzOtaeguiEtAl2016, WeideEtAl2018, DoucheWillcox2018, WhitlamEtAl2018}
and genetic data
\citep{BadrEtAl2000, MolinaCanoEtAl2005, KilianEtAl2007, OzkanEtAl2011, IobBotigue2023}
in recent years has further challenged the short-gestation model to
explain the origins of plant domestication and agriculture in West Asia.

Similarly, the concept of a limited set of eight `founder crops'
\citep{ZoharyHopf1988} that were the first species cultivated,
domesticated and then spread as the basis of Neolithic agricultural
systems, is not supported by the latest evidence. Our previous analyses
of the composition of available archaeobotanical datasets shows that
these crops were of marginal importance during the Epipalaeolithic
period \citep{ArranzOtaeguiEtAl2018} and that Neolithic subsistence did
not rely either solely or primarily on the exploitation of these species
\citep{ArranzOtaeguiRoe2023}. Instead, multiple species of grasses,
legumes, fruits, nuts, and other plants were exploited over the Late
Pleistocene-Early Holocene transition in southwest Asia.

\subsection{Biogeography and agricultural
origins}\label{biogeography-and-agricultural-origins}

The study of the natural distribution of the progenitors of domesticated
crops has been a central part of discussions on the origins of
agriculture and plant domestication from the beginning.
\citet{Humboldt1807} acknowledged the importance of the natural
distribution of wild species to explain the origin and domestication of
crops like spelt and rye. \citet{Candolle1886} integrated the study of
plant ecology and biogeography and influenced \citet{Darwin1859}, who
later reflected in detail about the geographical distribution of plants
and species diversity. At that time, there was intense debate about
whether there were single or multiple ``centres of creation'' of
species. Researchers aimed to evaluate whether plant and animal species
emerged in the same locations where they were currently distributed.

After Darwin, this early interest in crop origins evolved into more
specific discussions about the ``centres of plant domestication''.
\citet{Vavilov1926} was among the first to seek to determine the number
of regions in which plants had been independently domesticated
\citep{Harris1990}. His main method was `differential phytogeography':
he classified the variation within a crop and established the regions of
maximum diversity, to locate the geographic regions in which crops
originated. Using this method, Vavilov suggested that there were at
least eight centres of origin. His work was later criticised by
\citet{Harlan1971}, who argued that `centres of origin' and `centres of
diversity' had to be separated. For Harlan, a `centre' was as an ``area
in which things originate and out of which things are dispersed''
\citep[p.~468]{Harlan1971}, and he suggested that three main centres of
origin of domesticated crops existed. He further indicated that
Vavilov's approach to the question was simplistic and that more data
proxies had to be considered (e.g.~archaeology, history, geology), an
approach more in the tradition of \citet{Candolle1886}. Indeed, the
inclusion of archaeobotany and genetics in the last decades, together
with the study of wild relative distributions has been fundamental in
characterising the origins of agriculture \citep{FullerColledge2008}. As
a result of modern interdisciplinary studies, the number of recognised
centres of plant domestication has increased considerably, from the
three centres suggested by Harlan in 1971 to the six to eight centres
argued for in the 1990s \citep{Smith1995} and up to as much as 24
potential centres reported in 2009
\citetext{\citealp{PuruggananFuller2009}; \citealp[see
also][]{Fuller2010}}.

Biogeographic research into the centres of origin and/or domestication
of crops has also long informed broader understanding of the process of
agricultural origins. In \emph{Man Makes Himself}, \citet{Childe1936}
correctly located the centre of origin of European agriculture in the
`Fertile Crescent' of West Asia (unlike for example \citet{Pumpelly1908}
before him). This was not based on the region's prehistoric
archaeological record, which at the time had only been cursorily
explored. Instead he was guided to the region by biogeographic work by
Vavilov and \citet{PeakeFleure1927}; only later was this prediction
validated by archaeological work on the Epipalaeolithic and Neolithic of
Palestine \citep{Boyd2018}. In subsequent decades, the search for more
precise origin zones of specific domestic plants relied on the
assumption that ``the locus of domestication of a wild plant would
presumably be within its area of original distribution in the wild
state''
\citetext{\citealp{Butzer1971}; \citealp[paraphrasing][]{Helbaek1959}}
-- and that this ``natural habitat'' has not changed significantly over
the last 12,000 years \citep{Butzer1971}.

Contemporary research on crop origins was pioneered by Harlan and
Zohary, who compared the current distribution of the wild progenitors of
domesticated plants in southwest Asia
\citep{HarlanZohary1966, Zohary1969, Zohary1973, ZoharyHopf1973} to the
rapidly-expanding archaeobotanical record
\citetext{\citealp{Harlan1971}; \citealp{Harlan1977}; \citealp[see
also][]{ZoharyHopf1988}; \citealp{HarlanZohary1966}}. They both were
interested in evaluating which were the wild ancestors of domesticated
crops and studying their natural distribution to understand their
domestication process
\citep{Zohary1969, Zohary1973, ZoharySpiegelRoy1975}. Indeed, the
natural distribution of the wild relatives of domestic plant species was
later used as a criterion to infer `pre-domestic cultivation' in the
archaeological record. For example, the presence of seeds of chickpea at
Jericho led \citet{Hopf1986} to interpret the remains as cultivars, as
the natural distribution of the wild form of chickpea was located
further to the north. The same rationale was applied to the einkorn
remains found at several Pre-Pottery Neolithic sites in the southern
Levant \citep{Hopf1969, Colledge2001}, as the wild progenitors of this
species was thought to be restricted to the northern Levantine area
\citep{HeunEtAl1997, ZoharyEtAl2012}. The same idea---presence of plants
outside their natural range---has been repeated in the literature more
recently by several other authors
\citep{TannoWillcox2006, WillcoxEtAl2008, HillmanEtAl2001}.

Despite the importance of biogeography in the development and validation
of hypotheses regarding the origins of agriculture, there have been few
studies of the wild range of specific crop progenitors or other relevant
plant species \citep[cf.~for domestic animals, e.g.][]{YeomansEtAl2017}.
Observations regarding translocation or range expansion must therefore
rely on a relatively rough and ahistoric notion of a species' `natural
distribution' -- that is, one based primarily on contemporary or
recent-historic occurrences. Yet we know there has been considerable
climatic and environmental change in West Asia since the terminal
Pleistocene \citep{JonesEtAl2019}, so it is very unlikely that these
ranges were in fact static. Reconstructions of broader environments have
attempted to trace their fluctuations through time, either for the
entire region
\citetext{\citealp[e.g.][]{VanZeistBottema1991}; \citealp[Hillman
in][]{MooreEtAl2000}} or parts of it \citep[e.g.][]{Cordova2007}, but
these are at the level of the vegetation zone rather than individual
species. They also invariably rely on what might be called `expert
interpolation' (where the author composes a map based on his or her own
knowledge of the relevant data) rather than an explicit modelling
process. This makes it difficult, if not impossible, for users of such
reconstructions to understand exactly how they were derived or what
could explain, for example, the significant discrepencies between the
predictions of different experts.

\subsection{Ecological niche modelling in
archaeology}\label{ecological-niche-modelling-in-archaeology}

Ecological niche modelling or species distribution modelling is widely
used by ecologists to mathematically predict the geographic distribution
of a species from a sample of occurrences
\citep{FranklinMiller2009, SilleroEtAl2021}. Essentially, it involves
combining records of where an organism has been observed with
environmental data (climate, topography, etc.) for those locations to
model the range of environmental values at which that species -- its
environmental niche. This model can then be used to predict the
potential niche of the organism in question either in the same or a
different environment. \citet{TownsendPetersonSoberon2012} suggests
reserving the term `species distribution modelling' for when the method
is used to recover the verifiable range of a species in a real and
existing environment, and using `ecological niche modelling' as the
broader term covering hypothetical or predictive applications -- a
convention we follow here when referring to predictive or `hindcast'
models of past niches. Within this overarching framework, ecological
niche modelling encompasses a wide range of applications and a variety
of potential environmental predictors, modelling approaches, and
methodologies, which we will not attempt to review here.

Ecological niche modelling has long been of interest to archaeologists
as both a means of exploring the biological niche of humans and for
reconstructing the past environments they inhabited
\citep{DavidPollyEronen2011, FranklinEtAl2015}. In the first sense, it
has been used most extensively to model the niche of humans and other
hominin species
\citep[e.g.][]{BenitoEtAl2017, YousefiEtAl2020, BanksEtAl2021, YaworskyEtAl2024a, YaworskyEtAl2024b, GuranEtAl2024},
especially in the Palaeolithic. This overlaps with what archaeologists
usually call generically `predictive modelling'
\citep{VerhagenWhitley2020}---or more precisely `site distribution
modelling'---which is essentially the same approach as (and often
borrows methodologies from) ecological niche modelling but applied to
the occurrence of archaeological sites. Here what is modelled is not
strictly a biological niche alone, but also aspects of human geography,
taphonomy, and archaeological visibility. These applications can be
distinguished from `palaeoecological niche modelling', where the object
of model remains, as in ecology, a non-human biological niche.

\citet{FranklinEtAl2015} review palaeoecological niche modelling and
advocate for its greater adoption in environmental archaeology. In an
early application to West Asia, \citet{ConollyEtAl2012} used the
occurrence of wild and domestic \emph{Bos} remains at prehistoric
archaeological sites to map the evolving niche of cattle over the
Pleistocene--Holocene transition. It has been used to model the
availability of fauna exploited by humans at wider scales
\citep[e.g.][]{deAndresHerreroEtAl2018, YaworskyEtAl2023} and, in a West
Asian context, of foraged plant resources in the landscape around the
Neolithic sites on the Konya Plain \citep{CollinsEtAl2018}. Modelling
the spread of crops has been another significant archaeological
application \citep[e.g.][]{KrzyzanskaEtAl2022, Krzyzanska2023}, though
not as yet applied to West Asia.

In the majority of studies to date (palaeo)ecological niche modelling
has been applied to archaeological data in an `inductive' fashion,
i.e.~faunal and botanical remains from ancient sites are used as the
occurrence dataset for training a model using either past or present
environmental data. However, both the zooarchaeological and
archaeobotanical records are sparse and subject to a complex array of
depositional, taphonomic and recovery biases factors that , many of
which are not fully understood and/or cannot be corrected for. This
means that while the archaeological attestation of the presence of a
species might generally be relied upon, it is highly unlikely that its
absence is representative of true past distributions.

The alternative approach is to train the model using contemporary
occurrence and environmental data and then use palaeoenvironmental data
to `hindcast' its predictions backwards in time. Like
\citet{FranklinEtAl2015}, we view the hindcasting approach as more
promising, because training datasets for both occurrences and
environment are far more readily available, complete and reliable for
the present than the past. There is some scepticism in the ecological
niche modelling literature about the ability of such models to make
accurate predictions in unknown environments \citep[like the
past,][]{FranklinEtAl2015}, but here the hindcasting approach also
presents an opportunity: it reserves archaeological occurrence data as
an independent dataset that can be used to assess the retrodictive
performance of the model. This possibily was suggested by
\citet{FranklinEtAl2015} but to our knowledge our study represents the
first attempt to actually do so.

The major practical limitation of the hindcasting approach is that it
relies on spatially explicit, high resolution palaeoenvironmental
surfaces with continuous coverage of the region and periods of interest.
Until recently, this has not been widely available for most
applications, which is perhaps why only a minority of studies use it
\citep[cf.][]{KrzyzanskaEtAl2022, YaworskyEtAl2023}. In this study, we
are able to take advantage of the increasing availability of high
resolution, global palaeoclimate data derived from simulation
experiments with general circulation models of climate
\citep{BrownEtAl2018, BrownEtAl2020, KargerEtAl2023}.

\section{Data and model}\label{data-and-model}

\begin{figure}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{paper_files/figure-pdf/fig-region-1.pdf}}

}

\caption{\label{fig-region}Map of the study region (West Asia, grey box)
with locations of Late Epipalaeolithic and Pre-Pottery Neolithic
archaeobotanical assemblages.}

\end{figure}%

The aim of our study was to model the biogeography of species relevant
to human subsistence economies in West Asia (excluding the Southern
Arabian peninsula, see Figure~\ref{fig-region}) during the
archaeological Late Epipalaeolithic (15--11.7 ka) and Pre-Pottery
Neolithic (11.7--8.3 ka) periods. Based on current understandings, we
assume that plant-based subsistence during this period was broad,
geographically- and temporally-varied, and reflects a gradual,
geographically decentred, and nonlinear tranistion to greater reliance
on cultivars (i.e.~agriculture, see Section~\ref{sec-bg}). Our starting
point was a list of 68 taxa (Table~\ref{tbl-results-summary}) comprising
the identified species observed at at least 3 Late
Epipalaeolithic/Pre-Pottery Neolithic sites, according to our previous
study of the regional archaeobotanical data
\citep{ArranzOtaeguiRoe2023}. This was based on dataset collated from
three previously published regional archaeobotanical databases: ADEMNES
\citep{ADEMNES}, ORIGINS \citep{ORIGINS}, and COMPAG
\citetext{\citealp{LucasFuller2018}; \citealp{FullerEtAl2018}; \citealp[based
on][]{ColledgeEtAl2004}; \citealp{ShennanConolly2007}}. We did not
attempt to distinguish between the source of the remains
\citep[cf.][]{WallaceEtAl2018}; archaeobotanical assemblages are subject
to a variety of preservational and recovery biases, so by no means were
all the species on our list consumed or even deliberately collected by
people. However, we assume that there presence at a site of human
settlement at least implies that they were part of the wider ecosystem
that supported habitation there.

The taxonomic identifications of archaeobotanical material given in our
source databases were previously controlled to ensure consistency
between sources and to remove taxa that cannot be reliably distinguished
\citep[for details see][]{ArranzOtaeguiRoe2023}. Taxonomic names were
then matched to the canonical form specified in the GBIF Backbone
Taxonomy \citep{GBIFSecretariat2023} so they could be related to modern
occurrences. Archaeologically-attested domestic species meeting our
inclusion criteria were substituted for their wild progenitors (where
different) when gathering occurrence data, since the domestic forms are
now widespread and presumably uninformative of the species original
niche.

\subsection{Occurrence data}\label{sec-occ-data}

Georeferenced occurrence data for West Eurasia between 0 and 60° of
latitude was obtained from the Global Biodiversity Information Facility
(GBIF) using via and the R package `rgbif'
\citep{ChamberlainBoettiger2017, rgbif}. The GBIF dataset
\citep{GBIFDataDownload} excluded fossil occurrences, recorded absences,
and records with missing or dubious coordinates. Although niche models
have reasonable predictive power even with small training samples
\citep{StockwellPeterson2002, HernandezEtAl2006, WiszEtAl2008}, we did
not attempt to model 3 taxa with less than 40 usable occurrences,
following recommendations for niche models generally and Random
Forest-based models specifically
\citep{StockwellPeterson2002, LuanEtAl2020}. Multiple records of the
same taxon at the same coordinate were discarded because they do not
impart information to the model. The resulting cleaned dataset used to
train our niche models comprises 3392920 occurrences from 4769
constituent datasets.

\begin{figure}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{paper_files/figure-pdf/fig-occ-map-1.pdf}}

}

\caption{\label{fig-occ-map}Georeferenced occurrence records from West
Eurasia used to train models (N=1412083). Inset, right: prediction
region (West Asia).}

\end{figure}%

GBIF is currently the best available general-purpose occurrence dataset
for the West Asia region, its coverage is uneven both geographically and
from species to species. The Southern Levant, and Israel specifically,
is significantly more densely sampled than other parts of West Asia
(Figure~\ref{fig-occ-map}).

Random Forest is a presence--absence approach to niche modelling and
therefore requires not just data on where a species is present, but
where it is definitely not present. However `absence' data is rarely
available because it requires exhaustive survey. In practice, most
applications of niche modelling are `presence-only' and, where absence
data is required (as for Random Forest), it is supplied as a random
background sample of points.\footnote{We use a background sample that is
  representative of the entire study area, as opposed to attempting to
  construct a `pseudo-absence' sample. See \citet{SilleroEtAl2021} for
  the distinction.} The purpose of this sample is to inform the model
about the nature of the underlying environment. The stochastic
generation process means that some of these points will overlap or fall
close to presences, so ensuring the model is not overly influenced by
background samples is critical to its predictive importance
\citep{ValaviEtAl2022}. Here we follow the advice of
\citet{BarbetMassinEtAl2012} for regression-based species distribution
models and use a large (≈10000) uniform sample of points from across the
land area of the study region. These points are then weighted equally
against the presences in the regression to produce a `balanced Random
Forest' \citep{ValaviEtAl2022}.

\subsection{Predictor data}\label{sec-predictors}

We modelled the occurrence of species as a function of 24 geospatial
predictor variables. These included:

\begin{itemize}
\tightlist
\item
  Sixteen `bioclimatic' variables derived from monthly temperature and
  precipitation values, following standard practice for species
  distribution models \citep{HijmansEtAl2005}. Contemporary bioclimatic
  predictor data for West Asia was extracted from the global CHELSA
  dataset \citep{KargerEtAl2017}, which predicts temperature and
  precipitation from downscaled general circulation model output at 1 km
  resolution.
\end{itemize}

\begin{itemize}
\tightlist
\item
  Terrain aspect and slope, which at high resolution perform well as
  proxies for solar radiation when modelling plant occurrence
  \citep{AustinVanNiel2011, LeempoelEtAl2015}; and the topographic
  wetness index (TWI), which serves as a proxy for soil moisture and is
  particularly important in modelling arid environments
  \citep{KopeckyCizkova2010, CamposEtAl2016, DiVirgilioEtAl2018}. All
  three were derived from the SRTM30+ digital elevation model using
  algorithms from WhiteboxTools \citep{Lindsay2016}.
\end{itemize}

\begin{itemize}
\tightlist
\item
  Edaphic data from SoilGrids \citep{HenglEtAl2014, HenglEtAl2017},
  which improves model performance for plants
  \citep{DubuisEtAl2013, ModEtAl2016, VelazcoEtAl2017}. Based on a
  recent assessment of the reliability of SoilGrids data for species
  distribution modelling \citep{MillerEtAl2024}, we used a subset of
  four variables relating to soil texture (clay, silt, sand) and pH at
  the surface (0-5 cm depth).
\end{itemize}

For hindcasting, we used reconstructed bioclimatic data for three key
climatologies generated from downscaled paleoclimate simulations from
the HadCM3 general circulation model
\citep{FordhamEtAl2017, BrownEtAl2018}: the Bølling--Allerød
(c.~14.7--12.9 ka), the Younger Dryas (c.~12.9--11.7 ka), and the Early
Holocene (11.7--8.3 ka). Terrain and soil predictors were held constant,
since reconstructions of these variables in the past are not available
at sufficient scale. It is unlikely that either macroscale topography or
soil characteristics have altered significantly over the period of time
considered here, so we assume that this does not degrade model
performance, and may in fact benefit it by providing `anchoring'
predictors that are independent of climate change.

For training, test predictions, and archaeological predictions predictor
data was left in its native projection and resolution. For hindcast
palaeodistributions, it was transformed to common equal-area projection
and resolution of 5 km.

\subsection{Random Forest}\label{random-forest}

Ecological niche modelling is a classification problem that can be
approached with a wide range of statistical methods. A substantial
literature exists on the relatively performance of these approaches and
their respective parameterisations \citep[reviewed
in][]{ValaviEtAl2022}. Random Forest, a widely-used machine learning
algorithm, is amongst the best performing methods for presence-only
species distribution models, providing it is appropriately parameterised
to account for the class imbalance between presence and background
samples \citep{ValaviEtAl2021, ValaviEtAl2022}. For our application, it
also has the advantage of requiring little to no manual parameter tuning
to achieve good predictive results, which makes it easier to model a
larger numbers of taxa.

For each taxon we trained a classification model to predict occurrence
(presence or absence/background) based on up to 24 predictor variables
(Section~\ref{sec-predictors}). Highly correlated (Pearson's
\emph{r}\textgreater0.7) predictors were removed on a taxon-by-taxon
basis, to mitigate issues of overfitting due to colinearity
\citep{DormannEtAl2013}, as were redundant predictors with zero
variance. We used the Random Forest algorithm implemented in the R
package `ranger' \citep{WrightZiegler2017} and the `tidymodels'
\citep{tidymodels} framework for data preprocessing and model selection.
To avoid overfitting, we follow \citet{ValaviEtAl2021} in their
recommended hyperparameters and use of down-sampling to balance presence
and background samples. Models for each taxon were fit independently,
with redundant zero-variance predictors excluded, and assessed based on
balanced training (¾) and test (¼) partitions.

\section{Model assessment}\label{model-assessment}

We trained Random Forest models for 65 taxa using contemporary
occurrence data from GBIF, a random sample of background points, and the
predictor variables described in Section~\ref{sec-predictors}.
Substituting the ``current'' climate predictors for those derived from
palaeoclimatic simulations \citep{BrownEtAl2018}, we could then generate
hindcast predictions for reconstructed past environments in 4 key
climate periods -- a total of 260 modelled palaeodistributions.
Predicted distributions for individual taxa are presented in the
appendix and accompanying material.

\begin{figure}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{paper_files/figure-pdf/fig-model-sensitivity-vs-n_occ-1.pdf}}

}

\caption{\label{fig-model-sensitivity-vs-n_occ}Model sensitivity by
number of training occurrences}

\end{figure}%

We assessed the predictive performance of the fitted niche models in the
contemporary environment based on the reserved test partition
(Table~\ref{tbl-results-summary}). Model accuracy (proportion of
correctly classified presence and background samples) ranged between
93\% and 100\%, with an average of 98\%. Sensitivity (proportion of
correctly classified presence samples) ranged between 11\% and 100\%,
with an average of 83\%. The area under the models' receiver operating
characteristic curves (ROC-AUC) was on average 0.986±0.017. Model
sensitivity is loosely correlated with the number of occurrences
available for training (Figure~\ref{fig-model-sensitivity-vs-n_occ}),
with the worst-performing models all having less than 1000 recorded
occurrences: \emph{Triticum durum}, \emph{Euclidium syriacum},
\emph{Triticum turgidum}, \emph{Aegilops crassa}, and \emph{Vicia
orientalis}. Test metrics and ROC curves for the individual models are
included in the appendix.

The ability of the hindcast models to predict the occurrence of specific
species at archaeological sites is worse, with just 8\% of presences in
archaeobotanical assemblages successfully predicted at a threshold of
0.5. Model sensitivity (proportion of corrected predicted presenses) in
relation to the archaeological data is on average 0.06±0.12
(Table~\ref{tbl-results-summary}). A full assessment of the hindcasting
performance of the individual models can be found in the appendix.

In summary, the majority of our models perform poorly as simple binary
classifiers with a threshold of \texttt{P(present)\textgreater{}0.5}.
However, this is an arbitrary threshold; lowering it can improve
sensitivity (proportion of occurrences correctly predicted) at the cost
of specificity (proportion of non-occurrences incorrectly predicted).
There are several approaches to selecting informative thresholds for
binary classification in the ENM literature
\citep{LiuEtAl2005, LiuEtAl2013, LiuEtAl2015}, with most aiming for an
optimum compromise between sensitivity and specificity. We were unable
to find a thresholding technique that significantly improved the
hindcast models' performance as binary classifiers. By definition,
binarizing model output also discards information from the underlying
probabilistic prediction, so we have opted to present results
probabilistically as far as possible. In principle the archaeological
occurrences themselves could be used to `calibrate' the hindcast models
and select an optimum threshold, but there aren't enough archaeological
attestations of most of the modelled species for this to be practicable
at this time.

\begingroup
\setlength\LTleft{0\linewidth}
\setlength\LTright{0\linewidth}\fontsize{8.2pt}{9.9pt}\selectfont
\setlength{\LTpost}{0mm}

\begin{longtable}{@{\extracolsep{\fill}}lrrrrrr}

\caption{\label{tbl-results-summary}Summary of species modelled}

\tabularnewline

\toprule
 & \multicolumn{2}{c}{Occurrences} & \multicolumn{4}{c}{Model} \\ 
\cmidrule(lr){2-3} \cmidrule(lr){4-7}
Taxon & Arch. & Cur. & Acc. & ROC-AUC & Sens. cur. & Sens. arch. \\ 
\midrule\addlinespace[2.5pt]
{\itshape Aegilops crassa} & 4 & 223 & 0.98 & 0.98 & 0.26 & 0.00 \\ 
{\itshape Aegilops speltoides} & 0 & 448 & 0.99 & 0.99 & 0.77 & — \\ 
{\itshape Aegilops tauschii} & 0 & 1257 & 0.97 & 0.99 & 0.84 & — \\ 
{\itshape Aizoanthemopsis hispanica}\textsuperscript{\textit{1}} & 4 & 762 & 0.99 & 1.00 & 0.87 & 0.50 \\ 
{\itshape Ammi majus} & 4 & 10337 & 0.97 & 0.99 & 0.98 & 0.00 \\ 
{\itshape Androsace maxima} & 17 & 4946 & 0.94 & 0.98 & 0.90 & 0.00 \\ 
{\itshape Arenaria serpyllifolia} & 3 & 101617 & 0.98 & 0.99 & 1.00 & 0.00 \\ 
{\itshape Arnebia decumbens} & 21 & 617 & 0.97 & 0.97 & 0.59 & 0.05 \\ 
{\itshape Arnebia linearifolia} & 13 & 239 & 1.00 & 1.00 & 0.83 & 0.00 \\ 
{\itshape Atriplex prostrata} & 4 & 80186 & 0.99 & 0.99 & 0.99 & 0.00 \\ 
{\itshape Avena sterilis} & 5 & 15906 & 0.96 & 0.99 & 0.98 & 0.40 \\ 
{\itshape Bassia arabica} & 4 & 262 & 1.00 & 0.99 & 0.88 & 0.00 \\ 
{\itshape Bolboschoenus glaucus} & 5 & 418 & 0.98 & 0.97 & 0.63 & 0.00 \\ 
{\itshape Bolboschoenus maritimus}\textsuperscript{\textit{2}} & 31 & 53813 & 0.98 & 1.00 & 0.99 & 0.00 \\ 
{\itshape Brachypodium distachyon} & 5 & 13800 & 0.98 & 0.99 & 0.99 & 0.20 \\ 
{\itshape Bromus sterilis} & 3 & 89764 & 0.99 & 0.99 & 1.00 & 0.00 \\ 
{\itshape Buglossoides arvensis} & 23 & 23004 & 0.95 & 0.98 & 0.98 & 0.09 \\ 
{\itshape Buglossoides tenuiflora} & 26 & 399 & 0.99 & 0.98 & 0.79 & 0.04 \\ 
{\itshape Capparis spinosa} & 4 & 8617 & 0.96 & 0.99 & 0.95 & 0.00 \\ 
{\itshape Carex divisa} & 9 & 7400 & 0.96 & 0.99 & 0.97 & 0.00 \\ 
{\itshape Chenopodium album} & 5 & 184412 & 0.99 & 0.99 & 1.00 & 0.00 \\ 
{\itshape Cicer reticulatum}\textsuperscript{\textit{3}} & 3 & 52 & 1.00 & 1.00 & 1.00 & 0.00 \\ 
{\itshape Citrullus colocynthis} & 4 & 1823 & 0.94 & 0.96 & 0.71 & 0.00 \\ 
{\itshape Crithopsis delileana} & 3 & 430 & 1.00 & 1.00 & 0.93 & 0.00 \\ 
{\itshape Euclidium syriacum} & 3 & 338 & 0.98 & 0.96 & 0.31 & 0.00 \\ 
{\itshape Ficus carica} & 8 & 64440 & 0.98 & 0.99 & 0.99 & 0.38 \\ 
{\itshape Fumaria densiflora} & 4 & 3704 & 0.96 & 0.99 & 0.95 & 0.00 \\ 
{\itshape Gypsophila elegans} & 3 & 1613 & 0.96 & 0.99 & 0.83 & 0.00 \\ 
{\itshape Gypsophila pilosa} & 5 & 348 & 0.99 & 0.99 & 0.69 & 0.00 \\ 
{\itshape Gypsophila vaccaria}\textsuperscript{\textit{4}} & 8 & 1177 & 0.95 & 0.96 & 0.69 & 0.00 \\ 
{\itshape Henrardia pubescens} & 3 & 1\textsuperscript{\textit{5}} & — & — & — & — \\ 
{\itshape Hordeum bulbosum} & 4 & 4134 & 0.97 & 0.99 & 0.94 & 0.00 \\ 
{\itshape Hordeum murinum} & 5 & 74770 & 0.98 & 0.99 & 1.00 & 0.00 \\ 
{\itshape Hordeum spontaneum}\textsuperscript{\textit{6}} & 76 & 3098 & 0.99 & 1.00 & 0.98 & 0.21 \\ 
{\itshape Lathyrus aphaca} & 4 & 18652 & 0.96 & 0.99 & 0.98 & 0.00 \\ 
{\itshape Lathyrus oleraceus}\textsuperscript{\textit{7}} & 8 & 819 & 0.98 & 0.99 & 0.78 & 0.00 \\ 
{\itshape Lathyrus sativus} & 4 & 2120 & 0.93 & 0.97 & 0.77 & 0.00 \\ 
{\itshape Lepidium perfoliatum} & 3 & 2764 & 0.94 & 0.98 & 0.84 & 0.00 \\ 
{\itshape Linum bienne}\textsuperscript{\textit{8}} & 14 & 12824 & 0.98 & 0.99 & 0.99 & 0.07 \\ 
{\itshape Lolium rigidum} & 5 & 10620 & 0.97 & 0.99 & 0.98 & 0.00 \\ 
{\itshape Lolium temulentum} & 3 & 4463 & 0.94 & 0.98 & 0.94 & 0.00 \\ 
{\itshape Medicago radiata} & 20 & 834 & 0.98 & 1.00 & 0.90 & 0.30 \\ 
{\itshape Phalaris paradoxa} & 3 & 3287 & 0.96 & 0.99 & 0.94 & 0.00 \\ 
{\itshape Phragmites australis} & 4 & 373066 & 0.99 & 0.99 & 1.00 & 0.00 \\ 
{\itshape Pistacia atlantica} & 6 & 2419 & 0.98 & 1.00 & 0.93 & 0.33 \\ 
{\itshape Poa bulbosa} & 5 & 30769 & 0.96 & 0.98 & 0.99 & 0.20 \\ 
{\itshape Polygonum arenarium}\textsuperscript{\textit{9}} & 7 & 25\textsuperscript{\textit{5}} & — & — & — & — \\ 
{\itshape Polygonum corrigioloides} & 6 & 12\textsuperscript{\textit{5}} & — & — & — & — \\ 
{\itshape Prosopis farcta} & 5 & 2512 & 0.99 & 1.00 & 0.96 & 0.00 \\ 
{\itshape Rumex pulcher} & 6 & 17606 & 0.97 & 0.99 & 0.99 & 0.00 \\ 
{\itshape Salsola kali} & 6 & 14701 & 0.98 & 1.00 & 0.99 & 0.00 \\ 
{\itshape Salvia absconditiflora}\textsuperscript{\textit{10}} & 3 & 128 & 1.00 & 1.00 & 0.81 & 0.00 \\ 
{\itshape Secale cereale} & 4 & 14813 & 0.95 & 0.98 & 0.98 & 0.00 \\ 
{\itshape Secale strictum}\textsuperscript{\textit{11}} & 3 & 162 & 0.99 & 0.98 & 0.62 & 0.00 \\ 
{\itshape Suaeda fruticosa} & 3 & 577 & 0.99 & 0.99 & 0.76 & 0.00 \\ 
{\itshape Taeniatherum caput-medusae} & 4 & 1969 & 0.97 & 0.99 & 0.90 & 0.00 \\ 
{\itshape Triticum aestivum}\textsuperscript{\textit{12}} & 4 & 217 & 0.99 & 0.96 & 0.50 & 0.00 \\ 
{\itshape Triticum durum} & 3 & 65 & 0.99 & 0.95 & 0.11 & 0.00 \\ 
{\itshape Triticum monococcum}\textsuperscript{\textit{13}} & 47 & 870 & 0.97 & 0.99 & 0.80 & 0.04 \\ 
{\itshape Triticum turgidum}\textsuperscript{\textit{14}} & 53 & 345 & 0.98 & 0.91 & 0.25 & 0.00 \\ 
{\itshape Triticum urartu} & 0 & 420 & 1.00 & 1.00 & 0.95 & — \\ 
{\itshape Verbena officinalis} & 3 & 89655 & 0.99 & 0.99 & 1.00 & 0.00 \\ 
{\itshape Vicia ervilia} & 26 & 1924 & 0.95 & 0.98 & 0.84 & 0.27 \\ 
{\itshape Vicia faba} & 7 & 42644 & 0.97 & 0.99 & 0.99 & 0.29 \\ 
{\itshape Vicia narbonensis}\textsuperscript{\textit{15}} & 3 & 2826 & 0.95 & 0.99 & 0.90 & 0.33 \\ 
{\itshape Vicia orientalis}\textsuperscript{\textit{16}} & 16 & 41 & 1.00 & 0.97 & 0.14 & 0.00 \\ 
{\itshape Vitis sylvestris} & 3 & 63 & 1.00 & 1.00 & 0.50 & 0.00 \\ 
{\itshape Zygophyllum fabago} & 3 & 3554 & 0.97 & 0.99 & 0.94 & 0.00 \\ 
\bottomrule

\end{longtable}

\begin{minipage}{\linewidth}
\textsuperscript{\textit{1}}Including \emph{Aizoon hispanicum}\\
\textsuperscript{\textit{2}}Including \emph{Scirpus maritimus}\\
\textsuperscript{\textit{3}}Including \emph{Cicer arietinum}\\
\textsuperscript{\textit{4}}Including \emph{Vaccaria pyramidata}\\
\textsuperscript{\textit{5}}Excluded from modelling due to sample size\\
\textsuperscript{\textit{6}}Including \emph{Hordeum vulgare}\\
\textsuperscript{\textit{7}}Including \emph{Pisum sativum} and \emph{Pisum elatius}\\
\textsuperscript{\textit{8}}Including \emph{Linum usitatissimum}\\
\textsuperscript{\textit{9}}Including \emph{Polygonum venantianum}\\
\textsuperscript{\textit{10}}Including \emph{Salvia cryptantha}\\
\textsuperscript{\textit{11}}Including \emph{Secale montanum}\\
\textsuperscript{\textit{12}}Including \emph{Triticum spelta} and \emph{Triticum aestivocompactum}\\
\textsuperscript{\textit{13}}Including \emph{Triticum boeoticum}\\
\textsuperscript{\textit{14}}Including \emph{Triticum aestivum}, \emph{Triticum dicoccum}, and \emph{Triticum dicoccoides}\\
\textsuperscript{\textit{15}}Including \emph{Vicia narbonense}\\
\textsuperscript{\textit{16}}Including \emph{Lens culinaris} and \emph{Lens orientalis}\\
\end{minipage}
\endgroup

\section{Discussion}\label{discussion}

\subsection{Reduction in potential niche sizes over the
Pleistocene/Holocene
boundary}\label{reduction-in-potential-niche-sizes-over-the-pleistoceneholocene-boundary}

\begin{figure}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{paper_files/figure-pdf/fig-pred-range-density-1.pdf}}

}

\caption{\label{fig-pred-range-density}Distribution of predicted species
niche size by period (logarithmic scale). Dashed lines indicate the
median niche size.}

\end{figure}%

Our reconstructed palaeodistributions (shown in full in the appendix)
indicate that the majority of species had significantly different
potential geographic niches in the Late Pleistocene/Early Holocene
compared today. 55 of 65 species are predicted reduced niches in the
past; 52 of more than 10\% or more. Though the magnitude of the change
in potential niche size from prehistory to the present likely reflects a
degree of overfitting in the model (discussed further in
Section~\ref{sec-discuss-hindcasting}), fluctuations in modelled niche
size between the Bølling-Allerød (14.7--12.9 ka), Younger Dryas
(12.9--11.7 ka), and Early Holocene (11.7--8.3 ka) are more directly
comparable (Figure~\ref{fig-pred-range-density}). The average potential
niche size of modelled species was 35\% smaller in the Early Holocene
compared to the Bølling--Allerød, and 22\% smaller during the Younger
Dryas. This perhaps indicates that although this period is considered
one of climatatic amelioration globally \citep{JonesEtAl2019}, the
colder conditions of the Pleistocene may have supported more extensive
plant-based economies in West Asia specifically.

\begin{figure}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{paper_files/figure-pdf/fig-palaeodist-sum-1.pdf}}

}

\caption{\label{fig-palaeodist-sum}Predicted species richness (sum of
predicted ranges) by period}

\end{figure}%

Many taxa that occur (or are predicted to occur) across the `hilly
flanks' today---including most crop progenitors---are reconstructed to
have had a significantly more restricted distribution in the terminal
Pleistocene/Early Holocene (Figure~\ref{fig-palaeodist-sum}). These
include \emph{Ficus carica} (fig); \emph{Hordeum} spp. (wild barleys);
\emph{Lathyrus aphaca} and \emph{L. sativus} (both marginally edible
legumes); \emph{Triticum aestivum compactum} (in the N. Levant),
\emph{T. monococcum aegilopoides}, \emph{T. durum}, and \emph{Triticum
urartu} (but not the other wheat progenitor, \emph{T. turgidum dicoccum}
-- see Section~\ref{sec-discuss-crops}); \emph{Aegilops speltoides}, but
not \emph{Aegilops tauschii} (goatgrasses); and \emph{Vicia} spp.
(vetches), including \emph{Vicia faba} (broad beans). Most of Anatolia,
Northern Mesopotamia, and the Zagros Mountains in particular disappear
from the predicted niches of these species, leaving the Levant and to a
lesser extent the Aegean and Cyprus as refugia.

Our results for the Levant are consistent with the current understanding
of this region as developing early intensive foraging economies
\citep[the Natufian culture,][]{BarYosef1998} and as a centre of origin
of agriculture \citep{Zeder2011a}. Within the Levant, many species show
moderate reductions in potential niche size over the
Pleistocene/Holocene boundary, retreating from the Badia/transjordan
region (e.g.~barley, fig. Figure~\ref{fig-palaeodist-barley-levant}).

Loss of the Northern Mesopotamia--Anatolia region from the predicted
potential niches of crop progenitors is interesting in light of the
`golden triangle' hypothesis
\citep{Lev-YadunEtAl2000, KozlowskiAurenche2005, AbboEtAl2010}, which
puts this region at the centre of the development of agriculture and
plant domestication. Multiple lines of archaeological evidence have
emerged that point away from this hypothesis and towards a more
geographically diverse origin
\citep{Asouti2006, FullerEtAl2011, ArranzOtaeguiEtAl2016}, and our
reconstructions are also consistent with the late arrival of intensive
plant-based foraging economies in this region (cf.~the Natufian of the
Levant).

The near-absence of the Zagros in any predicted niches is also
surprising, given mounting evidence that animal domestication took place
just as early in the eastern \emph{Mashriq} as it did in the west
\citep{Zeder2024}. We consider that the most likely explanation for this
is that our flora does not include the species that were most important
to plant subsistence in the east. Archaeobotanical data on Neolithic
sites in the Zagros is limited (compared to the Levant in particular)
due to a hiatus in field research there from the 1980s to early 2010s
\citep{Zeder2024}. Recent research
\citep{RiehlEtAl2013, WeideEtAl2017, WeideEtAl2018, WhitlamEtAl2018, GonzalezCarreteroEtAl2023}
indicates that plant subsistence in this region was based on a distinct
set of species than that of the Levant and Anatolia.

Cyprus and the Aegean are not conventionally considered part of the
primary zone of domestication but rather amongst the first regions that
acquired agriculture from West Asia. Our analysis complicates this
picture, as it indicates that the wild distribution of many crop
progenitors included these regions. Early examples of several
domesticates are recorded at sites on Cyprus, Western Anatolia and
Greece \citep{ArranzOtaeguiRoe2023}, and the Aegean region was probably
connected to West Asia by a land bridge via Anatolia until the Early
Holocene \citep{AksuHiscott2022}. Were these area part of the same
broader `interaction sphere' that produced Neolithic agriculture in West
Asia?

Exceptions to the dominant trend of niche size reduction include
\emph{Cicer reticulatum} (wild chickpea), which has a relatively stable
potential niche centered on Northern Mesopotamia; and \emph{Triticum
turgidum dicoccum} (wild emmer wheat), which is predicted to occur in
two limited zones centered around the Black Sea Coast of Anatolia and
the Palmyra basin. In the latter case, neither of these areas are part
of the predicted (by our model) modern distribution of wild emmer, which
is centered around the Caucasus and Northern Mesopotamia. But it would
be consistent with archaeological evidence for early cultivation at
sites in the Upper Euphrates \citep{Willcox2024}.

\subsection{Biogeography of crop progenitors}\label{sec-discuss-crops}

\begin{figure}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{paper_files/figure-pdf/fig-palaeodist-chickpea-rye-1.pdf}}

}

\caption{\label{fig-palaeodist-chickpea-rye}Predicted palaeodistribution
of wild chickpea and rye in the Early Holocene (11.7--8.2 ka)}

\end{figure}%

Almost all the cereal and legume crop progenitors we modelled are
predicted to have only been found in the Levant during the terminal
Pleistocene and Early Holocene (see appendix). Part of this may be do
with the fact that both our initial flora and training occurrence
dataset have a strong bias towards the southern Levant, but the modelled
\emph{current} potential niche of these plants do tend to include
Anatolia and the Zagros, so this cannot be the only factor. It has
previously been proposed that the Southern Levant was an ``important
glacial refuge area for wild cereals'' \citep{RobertsEtAl2018}, which
accords with these results.

One notable exception is the wild ancestor of chickpea (\emph{Cicer
reticulatum}), which is predicted to have a distribution centered on
Northern Mesopotamia but encompassing much of the the `hilly flanks'
(except the southern Levant, Figure~\ref{fig-palaeodist-chickpea-rye}).
Another is rye (\emph{Secale cereale}), which is inferred to be
primarily Anatolian (Figure~\ref{fig-palaeodist-chickpea-rye}). This is
perhaps relevant to rye's unusual domestication history, as a crop of
West Asian origin that was intensively exploited
\citep{HillmanEtAl2001, DoucheWillcox2018} but apparently not first
cultivated until much later than the `founder crops', in Europe
\citep{SchreiberEtAl2021}.

\begin{figure}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{paper_files/figure-pdf/fig-palaeodist-locals-1.pdf}}

}

\caption{\label{fig-palaeodist-locals}Predicted palaeodistribution of
flax, pistachio and clubrush in the Early Holocene (11.7--8.2 ka)}

\end{figure}%

Flax (\emph{Linum bienne}) is predicted to have had a highly
concentrated distribution in Cyprus and along the Mediterranean coast of
the southern Levant. This is consistent with its low ubiquity in
archaeobotanical assemblages \citep{ArranzOtaeguiRoe2023}, despite
conventionally being considered a `founder crop', and presumably implies
that its domestication was similarly geographically constrained. It is
the only unambiguous crop progenitor with such a restricted potential
niche, though pistachio (\emph{Pistacia atlantica}) and clubrush
(\emph{Bolboschoenus maritimus}) are similarly constrained to the
Mediterranean coast (and North Africa, in the case of pistachio). This
is despite the fact that they are well-attested in the archaeotanical
record from across West Asia.

\begin{figure}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{paper_files/figure-pdf/fig-palaeodist-barley-levant-1.pdf}}

}

\caption{\label{fig-palaeodist-barley-levant}Predicted
palaeodistribution of wild barleys}

\end{figure}%

Wild barley (\emph{Hordeum spontaneum}) its relatives show a contraction
of their predicted potential niche from the Pleistocene to the Holocene,
concurrent with it being brought into cultivation
(Figure~\ref{fig-palaeodist-barley-levant}). It also sees a marked
decline in the archaeobotanical record from the Early PPNA/Early PPNB
(where it was amongst the most common taxa) to the Late PPNB and Late
Neolithic \citep{ArranzOtaeguiRoe2023}. Pistachio (\emph{Pistacia
atlantica}) shows similar trends, but it is less certain that this
species was managed in the Neolithic. Conversely, the various wild wheat
species native to West Asia show almost no response to
Pleistocene/Holocene climate change, even within the Levant, and in the
archaeobotanical record wheat displays the opposite trend to barley and
pistachio -- becoming gradually more abundant through the course of the
Neolithic and dominant by its end \citep{ArranzOtaeguiRoe2023}.

\begin{figure}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{paper_files/figure-pdf/fig-palaeodist-bread-1.pdf}}

}

\caption{\label{fig-palaeodist-bread}Combined predicted
palaeodistributions of bread wheat progenitors}

\end{figure}%

Bread wheat (\emph{Triticum aestivum}), the most common wheat cultivar
today, has a complex ancestry that involves two recent hybridisation
events \citep{LevyFeldman2022}: most recently between domestic emmer
(\emph{Triticum turgidum dicoccum}) and a goatgrass (\emph{Aegilops
tauschii}) c.~9 ka, and before that, in emmer, between wild red einkorn
(\emph{T. urartu}) and another goatgrass (\emph{Aegilops speltoides}).
The predicted potential niches of the two recent progenitors---\emph{T.
turgidum dicoccum} and \emph{A. tauschii}---have relateively limited
areas of overlap. These include the Syrian Jazira
(Figure~\ref{fig-palaeodist-bread}), close to the only two sites in our
archaeobotanical database where bread wheat (\emph{T. aestivum})
co-occurs with both of the recent progenitors (\emph{A. tauschii} and
\emph{T. turgidum dicoccum}): Tell Abu Hureyra and El Kowm II
\citep{ArranzOtaeguiRoe2023}.\footnote{We are grateful to Benjamin Nowak
  for pointing this out to us.} Taken together this suggests an origin
of domesticated bread wheat, otherwise only loosely geographically
constrained to the Levant--Upper Euphrates corridor
\citep{LevyFeldman2022}, in this vicinity.

\subsection{Disagreement between hindcast models and archaeobotanical
composition}\label{sec-discuss-hindcasting}

In general there is middling to low agreement between our hindcast
predictions of species niches and observed archaeological occurrences of
the same species \citep{tlb-results-summary}.

\begin{figure}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{paper_files/figure-pdf/fig-archaeo-sensitivity-1.pdf}}

}

\caption{\label{fig-archaeo-sensitivity}Summary of hindcasting
performance of models in relation to model performance and available
archaeological test data}

\end{figure}%

The hindcasting performance of models is not strongly correlated with
either the number of occurrences in the training dataset, the
performance of the model on test data, or the number of attested
archaeological occurrences (Figure~\ref{fig-archaeo-sensitivity}). This
indicates against a single, simple explanation for the discrepancy,
i.e.~that one or the other signal is `wrong'.

The core issue is that the two sets of evidence discussed
here---modelled palaeodistributions and archaeologically-attested
palaeooccurrences---are signals of related but distinct phenomena. The
hindcast niche models are projections of the species' present realised
niche into a reconstructed past climatology. The archaeobotanical record
is an incomplete sample of a past realised niche passed through various
anthropic and taphonomic filters. Neither can be assumed to be fully
accurate representations of the ultimate aim of this study, which is the
\emph{fundamental} niche of the species in relation to reconstructable
climate and environmental factors -- though both carry information about
it.

For each specific species there are a variety of potential explanations
for why the signals differ, but without further lines of evidence we are
not in a position to distinguish them. In general, our models probably
underestimate the fundamental niche or potential distribution of the
target taxon for a variety of reasons:

\begin{itemize}
\tightlist
\item
  The inherent tendency of machine learning models to overfit to the
  training scenario, despite our efforts to mitigate this;
\item
  Uneven spatial sampling intensity of the GBIF occurrence dataset (see
  Section~\ref{sec-occ-data}), meaning that some parts of the species'
  niche are probably either over- and under-represented in the training
  data;
\item
  Use of fairly coarse (2000 year) time-averaged climate slices, which
  capture a general trend but are probably unrepresentative of the
  environment around sites at the specific time at which they were
  occupied.
\end{itemize}

At the same time, there are several reasons to believe that the
archaeobotanical data (at least as it has been compiled for this study)
\emph{overestimates} the fundamental niche:

\begin{itemize}
\tightlist
\item
  Archaeological deposits are also time-averaged signals, in some cases
  over quite large spans of time;
\item
  The chronology attached to our archaeobotanical data is imprecise,
  e.g.~primarily based on point estimates of radiocarbon dates
  \citep{Michczynski2007}, producing a further averaging effect;
\item
  Archaeological occurences could be the result of human transplanting,
  range expansion and/or transport, especially considering this is the
  period where agriculture began to be practiced;
\item
  Most archaeobotanical samples in our archaeological test dataset were
  small, so individual errors and misidentifications have a
  correspondingly large impact on quantification of model performance.
\end{itemize}

At the same time, we cannot rule out more substantive reasons for the
discrepency between predicted and observed archaeological occurrences.
The niches of the modelled species could have changed since the Early
Holocene, which would not be captured in a model trained purely on
modern specimens. Human economic choices---mobility, foraging
strategies, cultivation, etc.---could also produce archaeobotanical
assemblages whose composition depart significantly from that of the
surrounding local flora.

What are the implications of this assessment? In the immediate term, the
modelled palaeodistributions presented here should be taken as
conservative estimates -- a minimal likely potential niche of the
species. Closer inspection of discrepencies between the modelled and
attested distributions of particular species in relation to specific
environmental, taphonomic and cultural factors could yield further
insights. Future applications of hindcast palaeoecological niche models
in archaeology could refine the methodology in several ways: using more
finely resolved palaeoclimate sequences \citep[e.g.][]{KargerEtAl2023};
building more refined archaeological chronologies; performing
hyperparameter tuning to further mitigate overfitting;\footnote{We did
  not perform hyperparameter tuning in this study due to computational
  limitations.} and compiling expanded archaeological datasets that
could be use to calibrate binary classification.

\section{Conclusion}\label{conclusion}

We present the first continuous, spatially explicit models of the
palaeodistributions of 65 plant species found regularly in association
with Late Epipalaeolithic and Early Neolithic sites in West Asia. This
deductive approach---modelling the niche of a species based on its
occurrence in relation to environmental factors today, and using this
together with palaeoclimate simulations to infer its past
distribution---represents a new line of evidence on the archaeoecology
of the world's first agricultural societies. It provides a complementary
picture to that gleaned from environmental archaeology and
climatological archives because it is independent of the taphonomic,
anthropic, and recovery-related processes that affect these records.

The modelled palaeodistributions of the species presented here (see
appendix) represent plausible minimum potential niche under the average
climatic conditions of the Bølling-Allerød, Younger Dryas, and Early
Holocene periods. The models' generally high performance as assessed
against independent contemporary test datasets lend confidence to these
predictions. Their application as a binary predictor of archaeological
presence or absence is significantly less promising, which likely
reflects a combination of methodological limitations of our modelling
approach, the incomplete and coarsely temporally-resolved nature of the
archaeobotanical test dataset, and genuine discrepencies between the
species' fundamental niche and its occurrence in the archaeological
record. Whether on the broad scale (e.g.~the restricted geographic range
of most species compared to their attestation in the archaeological
record), or relating to specific species (i.e.~false positives and false
negatives), these discrepencies suggest several avenues for future
investigation.

Modelling a large number of species using machine learning, the
substantial occurrence datasets available for the present day, and a
hindcasting approach to past distributions also represents a significant
advance in the methodology of palaeoecological niche modelling. This
approach is enabled by the availability of high quality, global open
datasets in ecology \citep{GBIF2025, GBIFSecretariat2023}, earth science
\citep{SRTM}, and climatology \citep{KargerEtAl2017, BrownEtAl2018}.
However, several areas of methodological improvement are evident.
Notably, the state of open data availability in archaeology lags
conspiciously behind the fields mentioned above. Though we have
benefited from the relatively long tradition of compiling
archaeobotanical data in our region of study
\citep{ColledgeEtAl2004, ShennanConolly2007, ADEMNES, LucasFuller2018, FullerEtAl2018, WallaceEtAl2018, ORIGINS},
further development of open, comprehensive and up-to-date `backbone'
datasets on site locations and chronologies is needed to advance
archaeoecological modelling to the same level.

\section{Data availability}\label{data-availability}

The data and R code used to produce this study is archived with Zenodo
at \url{https://doi.org/10.5281/zenodo.14629984}. Modelled
palaeodistributions in raster format (.TIF) can be found in the same
repository.

\section{Acknowledgements}\label{acknowledgements}

We are grateful to Alex Weide and the second, anonymous reviewer of our
manuscript, whose suggestions have greatly improved this study.

This work was partially funded by the European Research Council
(ERC-2021-STG project 101039060, \emph{PalaeOrigins: Tracing the
Epipalaeolithic origins of plant management in southwest Asia}).

\renewcommand\refname{References}
\bibliography{references.bib,gbif.bib}

\end{document}